A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.

### 2020-01-10 paper.Rmd
adds abstract (#91) and various phrasing

1 = Monash University, Department of Econometrics and Business Statistics

2 = Australia Centre of Excellence for Mathematical and Statistical Frontiers

3 = University of California, Berkeley

@@

@@

To share data analysis data, there should be some minimal set of requirements. For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)). No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others. It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \@ref(cite).

@@

Out of these sections, the minimal files to provide, in order of most importance should be provided in one main directory folder:

@@

In the context of datasets, READMEs are particularly useful when there are no reliable standards. The best practice here is to have one README per dataset.

A README is often the first place people will go to learn more about anything in a folder - they are very common in softwar, and historically were includd so that the uppecase lettrs of README meant ti would be at the top of a directory. In the context of data, a README when there are no reliable standards. It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":

@@

Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table taken from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate]. This includes information on the variable the class (type) of the variable, and a longer description of the variable.

@@

Data dictionary tables should be placed in the README and presented as a table. Every data dictionary should also be provided in its raw form in the repository, for example, saved as a CSV, so that they don't become trapped in a README.

@@

The CC0 is a “public domain” license. Data with a CC0 license means the data owners waive all their rights to the work, and it now "owned" by the public. The data can be freely shared, copied, modified, and distributed, even for commercial purposes _without asking permission_. When using data with CC0, it is good practice to cite the original paper, but it is not required. If you wish to use the CC0, see @choose-cc0.  For a brief overview of the CC0, see @cc0-short, and for the full license, see @cc0-long.

@@

The README and data dictionary provides _human readable_ information on the data. To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata. Machine readable metadata also provide a structure that allow the data to be indexed and searched online, through services scuh as google datasets search. An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema]. For example, a dataset called "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), which has the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).

@@

This contains fields such as "path", which describe the file path to the data, and a schema with subfields name and type, for each variable. It also provides information for licensing and features such as line breaks, and delimiters, useful to ensure the data is correctly imported into a machine. It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format. Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a lightweight markup format (YAML) using Table Schema.

@@

The data used in the data analysis should be provided in a folder called `data`. Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, observations in rows, and only one value per cell.

@@

## Publishing data in a data repository

>>>>

1 = Monash University, Department of Econometrics and Business Statistics, Melbourne, Australia

2 = Australian Centre of Excellence for Mathematical and Statistical Frontiers (ACEMS)

3 = University of California, Berkeley, USA

@@

**Abstract**

Data makes science possible. Sharing data improves visibility, and makes the research process transparent. This increases trust in the work, and allows for independent reproduction of results. However, a large proportion of data from published research is often only available to the original authors. Despite the obvious benefits of sharing data, and scientists' advocating for the importance of sharing data, most advice on sharing data discusses its broader benefits, rather than the practical considerations of sharing. This paper provides practical, actionable advice on how to actually share data alongside research. The key message is sharing data falls on a continuum, and entering it should come with minimal barriers.

@@

To share data analysis data, there should be some minimal set of requirements. For example, the data should contain information on metadata, data dictionaries, the README, and data used in analysis (See Section \@ref(minimal)). No matter where data is submitted, there should ideally be a canonical data repository in one long term archive that links to others. It should also have an accession number, or Digital Object Identification (DOI) number, which allows for it to be cited (discussed further in \@ref(cite)).

@@

From these sections, the minimal files to provide in order of most importance are:

@@

In the context of datasets, READMEs are particularly useful when there are no reliable standards. The best practice here is to have one README per dataset. A README is often the first place people will go to learn more about anything in a folder - they are very common in softwar, and historically were includd so that the uppecase lettrs of README meant ti would be at the top of a directory. In the context of data, a README when there are no reliable standards. It is meant for someone to read and understand more about the data and contains the "who, what, when, where, why, & how":

@@

Table \@ref(tab:table-data-incarcerate) shows an example data dictionary table from the Tidy Tuesday repository on incarceration trends [@tidy-tuesday-incarcerate]. This includes information on the variable, its class (type), and a longer description.

@@

Data dictionaries should be placed in the README and presented as a table. Every data dictionary should also be provided in its raw form (e.g., a CSV) in the repository, so they aren't "trapped" in the README.

@@

The CC0 is a "public domain" license. Data with a CC0 license means the data owners waive all their rights to the work, and it now "owned" by the public. The data can be freely shared, copied, modified, and distributed, even for commercial purposes _without asking permission_. When using data with CC0, it is good practice to cite the original paper, but it is not required. If you wish to use the CC0, see @choose-cc0.  For a brief overview of the CC0, see @cc0-short, and for the full license, see @cc0-long.

@@

The README and data dictionary provides _human readable_ information on the data. To ensure data types are preserved - dates are dates, names are characters, and so on - there needs to be some form of _machine readable_ metadata. This provides a structure allowing the data to be indexed and searched online, through services scuh as google datasets search [@google-data-search]. An excellent standard for metadata is Table Schema written by frictionless data [@fritionless-table-schema]. For example, a dataset "demographics" with three variables is shown in Table \@ref(tab:table-machine-readable), and the Java Script Object Notation (JSON) equivalent in Figure \@ref(fig:json-code).

@@

This contains fields such as "path", describing the file path to the data, and a schema with subfields name and type, for each variable. It also provides information for licensing and features such as line breaks, and delimiters, useful to ensure the data is correctly imported into a machine. It is built on JSON, a lightweight, human-friendly, machine readable data-interchange format. Table schema is baked into formats such as csvy [@csvy], an extended `csv` format, which has additional front matter in a lightweight markup format (YAML) using Table Schema.

@@

The data used in the data analysis should be provided in a folder called `data`. Ideally, the data "analysis ready data" should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, observations in rows, and only one value per cell.

@@

## Publishing data in a data repository (need some help here)
